---
title: "Time series analysis on pollution measurement data in New York"
author: "Guillem Bagaria, Juan Sebastian Garc??a, Amy Zhang"
date: "4/2/2018"
output: html_document
---

We obtained the data from an old kaggle competition: https://www.kaggle.com/sogun3/uspollution/data. This is much more data than we needed as it includes many different indicators as well as hundreds of locations.

For the mean values of Carbon Monoxide, we will perform a volatility analysis. The objective of this is to do a basic volatility analysis using GARCH.

Setup, loading and data cleaning
```{r eval=FALSE, include=FALSE}
rm( list=ls() )

library(fGarch)
library(tseries)
library(moments)
library(dplyr)
library(xts)
library(data.table)
source('/Users/guillembp/Desktop/Financial Econometrics/code/Volatility Modelling/lib/utilities.R')
source('/Users/guillembp/Desktop/Financial Econometrics/code/Volatility Modelling/lib/tarch.R')
source('/Users/guillembp/Desktop/Financial Econometrics/code/Volatility Modelling/lib/egarch.R')

wd='/Users/guillembp/Desktop/Financial Econometrics/Project/'
setwd(wd)
pollution<- read.csv('/Users/guillembp/Desktop/Financial Econometrics/Project/data/pollution_us_2000_2016.csv',header = TRUE)
```

```{r}
# Data cleaning


pollution$Date.Local<-as.Date(pollution$Date.Local,format="%Y-%m-%d")

pollution <- pollution[pollution$City == "New York",]
pollution <- pollution[pollution$Site.Num == 133,]
pollution <- pollution[pollution$Date.Local>="2014-01-01",]
pollution <- pollution[!is.na(pollution$CO.Mean),]
pollution <- pollution[ seq(1, nrow(pollution), 2) ,]
# convert to dialy data ----
ts.data <- xts(pollution$CO.Mean,
               order.by = pollution$Date.Local)

dialy_<-period.apply(ts.data,endpoints(ts.data,"days"),max)

ind=index(dialy_)
values=as.vector(dialy_)
dialy_poll<-data.frame(y=values,date=ind)
dates_poll <- dialy_poll$date
CO_units_ <- dialy_poll$y
T=length(dates_poll)

```

Plot initial data
```{r}
dates <- as.Date(as.character(pollution[,9]),'%Y-%m-%d')

plot(pollution$CO.Mean, type = "l", xlab = 'Measurements from 2014 to 2016', ylab = 'Measured CO Mean values')
```

We notice the seasonality, that appears to increase values in the winter. As we're measuring CO concentration, this is expected as most of the CO will be coming from old home heaters, with insufficient draft, instead of transport.

The analysis is based on the returns, calculated for this data through the formula below.

```{r}
ret <- diff(log(CO_units_))*100

ret <- c(ret,0)

plot(ret, type = "l", xlab = 'Measurements from 2014 to 2016', ylab = 'Measured CO Mean values')
```

We observe the data as very noisy. Seasonality is now not in the value, which has become stationary, but is still reflected in the variance.

The data is now prepared to be analysed for volatility.

 Volatility Clustering
```{r}
myplot(ind, abs(ret), col='red' )
abline(h=0, lwd=2)


# ACF ret
ret.acf <- acf(ret , ylim=c(-0.3,1) , lwd=5 , xlim=c(0,25) , col='darkorange2')

# PACF ret
ret.acf <- pacf(ret , ylim=c(-0.3,1) , lwd=5 , xlim=c(0,25) , col='darkorange2')

```

The ACF shows that the 1st and 2nd lags are significant, while the PACF shows a slowly decreasing correlation. In order to caputer this we suggest a higher order GARCH.

 GARCH
```{r}
garch21 <- garch(ret, order = c(2,1))

summary(garch21)

sigma <- garch21$fitted.values[,1]

myplot( ind , sqrt(252)*sigma , col='red2' )

myplot( ind , ret , col='orange2' )
lines( ind , 1.96*sigma , col='blue2')
lines( ind , -1.96*sigma , col='blue2')
```

# Residuals
```{r}
z <- ret/sigma
myplot( ind , z , col='red2' )

par( mar=c(2,2,0.1,0.1) )
qqnorm(ret,col='tomato',main='')
qqline(ret,distribution = qnorm)

par( mar=c(2,2,0.1,0.1) )
qqnorm(z,col='tomato',main='')
qqline(z,distribution = qnorm)

par( mar=c(2,2,0.1,0.1) )
acf( abs(ret) , lag.max=100 , ylim=c(-0.5,1) , lwd=5 , xlim=c(0,25) , col='darkorange2')

par( mar=c(2,2,0.1,0.1) )
acf( abs(z[4:length(z)]) , lag.max=100 , ylim=c(-0.1,1) , lwd=5 , xlim=c(0,25) , col='darkorange2')

```

The residuals follow a normal distribution. However, there appears to be some outliers. It seems that for this data we should include one or more dummy variables to account for exogenous effects that produce the outlying observations.

```{r}
jarque.test(ret)

```

The high value of the Jarque-Bera test indicates Normality of the residuals and thus we can confidently use GARCH to model this variance.
